---
title: 'Effective R: LGBM'
author: "Ryuta Yoshimatsu"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message = FALSE}
library(data.table)
library(Matrix)
library(dplyr)
library(MLmetrics)
library(lightgbm)
library(ggplot2)
library(gridExtra)
library(grid)
library(graphics)
library(TTR)
library(forecast)
library(lubridate)
library(mltools)
library(data.table)
library(ggplotify)
library(gridBase)
library(tsibble)
library(fable)
```


```{r}
label <- paste("@ryutayoshimatsu / D ONE /", as.character(Sys.Date()), sep=" ")
```

Read the data. 

```{r, fig.width=15, fig.height=7.5, warning = FALSE}
df <- fread("fitted.csv")
df <- df[, delta_cases := c(0, diff(cases))][]
df <- df[, trend := c(0, diff(fitted))][]
df <- df %>% mutate(detrend_delta = delta_cases - trend)

p1 <- ggplot(df, aes(x=time)) + 
  geom_line(aes(y = cases, color="observed")) + 
  geom_line(aes(y = fitted, color="fitted")) + 
  ggtitle("Total Infected Cases") +
  scale_color_manual(name = "", values = c("observed" = "black", "fitted" = "darkred")) +
  annotate("text", x=2021.305, y=1.9e5, size=2.5, label=label, angle=90) + 
  annotate("text", x=2020.00, y=5.75e5, size=3, label='2020-01-01') + 
  annotate("text", x=2020.25, y=5.75e5, size=3, label='2020-04-01') + 
  annotate("text", x=2020.50, y=5.75e5, size=3, label='2020-07-01') + 
  annotate("text", x=2020.75, y=5.75e5, size=3, label='2020-10-01') + 
  annotate("text", x=2021.00, y=5.75e5, size=3, label='2021-01-01') + 
  annotate("text", x=2021.25, y=5.75e5, size=3, label='2021-04-01') + 
  xlab("") +
  coord_cartesian(xlim = c(2020, 2021.25), clip = "off") +
  theme_minimal()
  
p2 <- ggplot(df, aes(x=time)) + 
  geom_line(aes(y = delta_cases, color="observed")) + 
  geom_line(aes(y = trend, color="fitted")) + 
  ggtitle("New Infected Cases") +
  scale_color_manual(name = "", values = c("observed" = "black", "fitted" = "darkred"))+
  annotate("text", x=2021.305, y=3000, size=2.5, label=label, angle=90) + 
  annotate("text", x=2020.00, y=9.2e3, size=3, label='2020-01-01') + 
  annotate("text", x=2020.25, y=9.2e3, size=3, label='2020-04-01') + 
  annotate("text", x=2020.50, y=9.2e3, size=3, label='2020-07-01') + 
  annotate("text", x=2020.75, y=9.2e3, size=3, label='2020-10-01') + 
  annotate("text", x=2021.00, y=9.2e3, size=3, label='2021-01-01') + 
  annotate("text", x=2021.25, y=9.2e3, size=3, label='2021-04-01') + 
  coord_cartesian(xlim = c(2020, 2021.25), clip = "off") +
  theme_minimal()

grid.arrange(
  p1, p2,
  nrow = 2,
  bottom = textGrob(
    "",
    gp = gpar(fontface = 3, fontsize = 9),
    hjust = 1,
    x = 1
  )
)
```

```{r}
df
```


```{r, fig.width=15, fig.height=3.25, warning = FALSE}
ggplot(df, aes(x=time)) + 
  geom_line(aes(y = delta_cases), color="black") + 
  geom_line(aes(y = trend), color="darkred") + 
  ggtitle("New Infected Cases") +
  annotate("text", x=2021.305, y=3000, size=2.5, label=label, angle=90) + 
  annotate("text", x=2020.00, y=9.2e3, size=3, label='2020-01-01') + 
  annotate("text", x=2020.25, y=9.2e3, size=3, label='2020-04-01') + 
  annotate("text", x=2020.50, y=9.2e3, size=3, label='2020-07-01') + 
  annotate("text", x=2020.75, y=9.2e3, size=3, label='2020-10-01') + 
  annotate("text", x=2021.00, y=9.2e3, size=3, label='2021-01-01') + 
  annotate("text", x=2021.25, y=9.2e3, size=3, label='2021-04-01') + 
  coord_cartesian(xlim = c(2020, 2021.25), clip = "off") +
  theme_minimal()
```



```{r, fig.width=15, fig.height=3.25, warning = FALSE}
ggplot(df, aes(x=time)) + geom_line(aes(y = detrend_delta), color = "black") +  
  ggtitle("Detrended Daily Infected Cases") +
  xlab("Time") +
  annotate("text", x=2020.00, y=1.8e3, size=3, label='2020-01-01') + 
  annotate("text", x=2020.25, y=1.8e3, size=3, label='2020-04-01') + 
  annotate("text", x=2020.50, y=1.8e3, size=3, label='2020-07-01') + 
  annotate("text", x=2020.75, y=1.8e3, size=3, label='2020-10-01') + 
  annotate("text", x=2021.00, y=1.8e3, size=3, label='2021-01-01') + 
  annotate("text", x=2021.25, y=1.8e3, size=3, label='2021-04-01') + 
  annotate("text", x=2021.305, y=-1.5e3, size=2.5, label=label, angle=90) +
  coord_cartesian(xlim = c(2020, 2021.25), clip = "off") +
  theme_minimal()
```

1. Linear combination of three logistic curves does very well at fitting to the total infected cases and estimating the overall trend.
2. But it does not to capture the seasonality.

Decompose the detrended series into trend, seasonal and residual components.

```{r, fig.width=15, fig.height=7.5, warning = FALSE}
detrend_delta.ts <- msts(df$detrend_delta, seasonal.periods=c(7, 30.4167))
detrend_delta.ts %>% mstl() %>% autoplot(main='Additive Multi-Seasonality Decomposition', xlab='Month')
```

1. Detrended cases of newly infected incidents has two prominent seasonality components: weekly and monthly. 
2. Remainder still exhibits structures: volatility.

We perform regression with ARIMA errors including Fourier terms with base periodicity at 7 and 30.4167 as additional regressors. The smoothness of the seasonal pattern are controlled by K (the number of Fourier sin and cos pairs – the seasonal pattern is smoother for smaller values of K). We use AIC to find the optimal K.

```{r, fig.width=15, fig.height=5}
bestfit <- list(aicc=Inf)
for(i in 1:3)
  for(j in 1:5)
  {
    fit <- auto.arima(detrend_delta.ts, xreg=fourier(detrend_delta.ts, K=c(i,j)), stationary=FALSE, seasonal=FALSE)
    if(fit$aicc < bestfit$aicc)
      bestfit <- fit
    else break;
  }
summary(bestfit)
```

```{r}
# Number of days to forecast 
h <- sum(is.na(df$cases))
print(h)
df
```

```{r, fig.width=15, fig.height=3.25, warning=FALSE, message=FALSE}
detrend_delta.forecasts <- forecast(bestfit, xreg=fourier(detrend_delta.ts, K=c(3,1), h=h))

par(mfrow=c(1, 2))

plot(bestfit$x, col="black", main="Fit to the Observed Daily Infected Cases", xlab="Month", ylab="detrended_delta")
lines(fitted(bestfit), col="darkred")
legend(0.85, -1800, legend=c("observed", "fitted"), col=c("black", "darkred"), lty=1:1, cex=0.75)
grid(NULL, NULL)
plot.new()

vps <- baseViewports()
pushViewport(vps$figure)
vp1 <- plotViewport(c(1,1,1,1))
f <- autoplot(detrend_delta.forecasts) + 
  autolayer(detrend_delta.forecasts$mean, series="Forecasts") + 
  theme_minimal() +
  ylim(-2500, 2000) + 
  xlab('Month') +
  ylab('detrended_delta')

print(f, vp=vp1)
```

```{r}
seasonal <- append(data.frame(fitted(bestfit))[1:(length(fitted(bestfit))-h),], data.frame(detrend_delta.forecasts$mean)[1:h,])
df <- cbind(df, seasonal)
df <- df %>% mutate(predicted = trend + seasonal)
df$predicted[df$predicted < 0] <- 0
```

```{r, fig.width=15, fig.height=3.25, warning=FALSE}
ggplot(df, aes(x=time)) + 
  geom_line(aes(y=delta_cases, color="observed")) + 
  geom_line(data=tail(df, h), aes(x=time, y=predicted, color="predicted")) +
  ggtitle("Forecast Daily Infected Cases") +
  annotate("text", x=2021.305, y=3250, size = 2.5, label=label, angle=90) +
  annotate("text", x=2020., y=9400, size=3, label='2020-01-01') + 
  annotate("text", x=2020.25, y=9400, size=3, label='2020-04-01') + 
  annotate("text", x=2020.50, y=9400, size=3, label='2020-07-01') + 
  annotate("text", x=2020.75, y=9400, size=3, label='2020-10-01') + 
  annotate("text", x=2021.00, y=9400, size=3, label='2021-01-01') + 
  annotate("text", x=2021.25, y=9400, size=3, label='2021-04-01') +
  coord_cartesian(xlim = c(2020, 2021.25), clip = "off") +
  xlab("Time") +
  scale_color_manual(name = "", values = c("observed"="black", "predicted"="darkred")) +
  theme_minimal()
```

(Compare this prediction with currently available number of cases and uplift if smaller.)

```{r, warning=FALSE, message=FALSE}
estimated_R <- read.csv(url("https://raw.githubusercontent.com/covid-19-Re/dailyRe-Data/master/CHE-estimates.csv"))
estimated_R <- estimated_R %>% group_by(date) %>% summarise(mean_R = mean(median_R_mean))
estimated_R$date <- as.Date(estimated_R$date, "%Y-%m-%d")
estimated_R <- estimated_R[which(estimated_R$date >= as.Date('2020-02-25', "%Y-%m-%d")), ]
mean_R <- estimated_R$mean_R
for (i in (length(mean_R)+1):(length(df$time)))
  mean_R[i] <- NA
```

```{r}
df <- cbind(df, mean_R)
```

```{r, fig.width=15, fig.height=3.75, warning=FALSE}
coefficient <- 5
ggplot(df, aes(x=time)) +
  
  geom_line(aes(y=log(delta_cases), colour='delta_cases')) + 
  geom_line(aes(y=mean_R*coefficient, colour='R')) + 
  ggtitle("Daily Infected Cases vs. R") +
  annotate("text", x=2021.305, y=4.75, size = 2.5, label=label, angle=90) +
  annotate("text", x=2020.00, y=14, size=3, label='2020-01-01') + 
  annotate("text", x=2020.25, y=14, size=3, label='2020-04-01') + 
  annotate("text", x=2020.50, y=14, size=3, label='2020-07-01') + 
  annotate("text", x=2020.75, y=14, size=3, label='2020-10-01') + 
  annotate("text", x=2021.00, y=14, size=3, label='2021-01-01') + 
  annotate("text", x=2021.25, y=14, size=3, label='2021-04-01') +
  coord_cartesian(xlim = c(2020, 2021.25), clip = "off") +
  xlab("Time") +
  scale_y_continuous(name="log(delta_cases)", sec.axis = sec_axis(~.*1./coefficient, name="R")) +
  scale_color_manual(name = "", values = c("delta_cases" = "black", "R" = "darkred")) +
  geom_hline(yintercept=coefficient, linetype = 'dotted', color='black') +
  theme_minimal()
```

```{r}
correlation_df <- df %>% select(delta_cases, mean_R)
correlation_df$delta_cases <- log(correlation_df$delta_cases)
correlation_df <- correlation_df[-1,]
correlation_df <- correlation_df[complete.cases(correlation_df), ]
cat('correlation between number of cases and the log of effective R: ', cor(correlation_df$delta_cases, correlation_df$mean_R))
```

The correlation coefficient between R and new cases is -0.25, which suggests that there is no strong 'linear' correlation between the two variables. But, since R is the rate of exponential growth in the ODE of any transmission models, I would expect some degree of relationship between the log of new cases and R.  

```{r, fig.width=7.5, fig.height=7.5}
mean_R_ <- na.omit(df$mean_R)
lag.plot(mean_R_, lags=9, main="Scatter Lag Plots: mean_R")
```

Prepare the training data table for modeling.

```{r}
# Prepare training data table
X <- df %>% select(time, delta_cases, mean_R)
X <- X  %>% rename(cases = delta_cases, R = mean_R)
```

Function to create lagged and rolling window features.

```{r}
create_features <- function(dt) {
  
  # Add lag vectors: table must be sorted by date!
  R_lags <- c(1, 2, 3, 4, 5, 6, 7)
  R_lag_cols <- paste0("R_lag_", R_lags)
  dt[, (R_lag_cols) := shift(.SD, R_lags), .SDcols="R"]
  
  cases_lag <- c(1, 2, 3, 4, 5, 6, 7)
  cases_lag_cols <- paste0("cases_lag_", cases_lag)
  dt[, (cases_lag_cols) := shift(.SD, cases_lag), .SDcols="cases"]
  
  # Add rolling window vectors: table must be sorted by date!
  R_windows <- c(7)
  R_roll_cols <- paste0("R_rmean_", t(outer(R_lags, R_windows, paste, sep="_")))
  dt[, (R_roll_cols) := frollmean(.SD, R_windows, na.rm=TRUE), .SDcols=R_lag_cols] # Rolling features on lag_cols
  
  cases_windows <- c(7)
  cases_roll_cols <- paste0("cases_rmean_", t(outer(cases_lag, cases_windows, paste, sep="_")))
  dt[, (cases_roll_cols) := frollmean(.SD, cases_windows, na.rm=TRUE), .SDcols=cases_lag_cols] # Rolling features on lag_cols

  return(dt)
}
```

```{r}
X <- create_features(X)
X <- na.omit(X)
sample <- head(X)
sample
sample <- sample[, c(1, 3, 2, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31)]
write.table(sample, file = "sample.csv")
```

```{r}
# Labels for the training set
y <- X$R

# Drop columns "time" and "R"
X[, c("time", "R") := NULL]
```

Convert a data frame to a numeric matrix: return the matrix obtained by converting all the variables in a data frame to numeric mode and then binding them together as the columns of a matrix.

```{r}
X <- data.matrix(X)
```

Construct lgb dataset.

```{r}
# Split the training data set into train and eval:
#     train consists of data from "2020-02-25" to 14 days prior to the last record available in X
#     val  consists of data from the last 14 days of X

## Indexes for the training set
#idx <- c(1:(length(X$time)-14))

# walk-forward validation
X_walk <- X
y_walk <- y
```

We use Poisson regression (from generalize linear model family), which is suitable for counts. The model assumes the errors are Poission distributed and thus could capture a skew, discrete distribution, and the restriction to response variables to be non-negative is applied.

```{r, warning=FALSE}
start_date = 30              # 2020-04-01
forecast_window = 14         # forecast horizon
end_date = nrow(X)-forecast_window
count = 0 
rmse_evolution <- data.frame()

for (date in c(start_date:end_date))
{
  idx_train <- c(1:(date-1))
  idx_val <- c(date:(date+forecast_window-1))
  #idx_val <- date
  
  xtrain <- lgb.Dataset(X_walk[idx_train, ], label=y_walk[idx_train]) 
  xval <- lgb.Dataset(X_walk[idx_val,,drop=F], label=y_walk[idx_val])
  
  # Configure lgb hyper parameters 
  p <- list(objective = "poisson",  # Training parameter
            metric ="rmse",         # Training parameter
            force_row_wise = TRUE,  # Training parameter: force row-wise histogram building
            learning_rate = 0.075,  # Training parameter
            num_leaves = 34,        # Regularization parameter
            min_data = 10,          # Regularization parameter
            sub_feature = 0.8,      # Regularization parameter
            sub_row = 0.75,         # Regularization parameter
            bagging_freq = 1,       # Regularization parameter
            lambda_l2 = 0.1,        # Regularization parameter
            nthread = 2)            # Training parameter        
  
  model.lgb <- lgb.train(params = p,
                     data = xtrain,               # Training set
                     valids = list(val = xval),   # Validation set
                     nrounds = 500,               # Training parameter (max number of trees)
                     early_stopping_rounds = 100, # Training parameter (min number of trees to stop)
                     eval_freq = 50,              # Training parameter
                     verbose = -1)             
  
  if (count %% 10 == 0)
  {
    cat("Best rmse on the validation set at day", count, ":", model.lgb$best_score, "at", model.lgb$best_iter, "iteration", "\n")
  }
  
  metric <- data.frame(count+1, model.lgb$best_score)
  rmse_evolution <- rbind(rmse_evolution, metric)
  count <- count + 1
}

rmse_evolution <- rmse_evolution%>% rename(time = count...1, rmse = model.lgb.best_score)
```

```{r, fig.width=15, fig.height=7.5, warning = FALSE}
p1 <- ggplot(rmse_evolution, aes(x=time)) + geom_line(aes(y = rmse), color = "black") +  
  ggtitle("Walk-Forward Validation: RMSE for Multistep (14 Days) Prediction") +
  annotate("text", x=-91, y=0.25, size=3, label='2020-01-01') + 
  annotate("text", x=0, y=0.25, size=3, label='2020-04-01') + 
  annotate("text", x=91, y=0.25, size=3, label='2020-07-01') + 
  annotate("text", x=183, y=0.25, size=3, label='2020-10-01') + 
  annotate("text", x=275, y=0.25, size=3, label='2021-01-01') + 
  annotate("text", x=365, y=0.25, size=3, label='2021-04-01') + 
  annotate("text", x=385, y=0.085, size=2.5, label=label, angle=90) +
  coord_cartesian(xlim = c(-91, 366), ylim = c(0, 0.25), clip = "off") +
  xlab("Day") +
  ylab("RMSE")

p2 <- ggplot(df, aes(x=time)) + 
  geom_line(aes(y = delta_cases)) + 
  ggtitle("New Infected Cases") +
  xlab("Time") +
  geom_line(aes(y = trend), color="darkred") + 
  annotate("text", x=2021.305, y=3000, size = 2.5, label=label, angle=90) +
  annotate("text", x=2020., y=9e3, size=3, label='2020-01-01') + 
  annotate("text", x=2020.25, y=9e3, size=3, label='2020-04-01') + 
  annotate("text", x=2020.50, y=9e3, size=3, label='2020-07-01') + 
  annotate("text", x=2020.75, y=9e3, size=3, label='2020-10-01') + 
  annotate("text", x=2021.00, y=9e3, size=3, label='2021-01-01') + 
  annotate("text", x=2021.25, y=9e3, size=3, label='2021-04-01') + 
  coord_cartesian(xlim = c(2020, 2021.25), clip = "off")

grid.arrange(
  p1, p2,
  nrow = 2,
  bottom = textGrob(
    "",
    gp = gpar(fontface=3, fontsize=9),
    hjust = 1,
    x = 1
  )
)
```

```{r}
imp <- lgb.importance(model.lgb)
imp[order(-Gain)
    ][1:length(imp$Feature), ggplot(.SD, aes(reorder(Feature, Gain), Gain)) +
        geom_col(fill = "darkblue") +
        xlab("Feature") +
        coord_flip() +
        theme_minimal()]
```

As we are using lag features we have to forecast day by day in order to use the latest predictions for the current day.

```{r}
# Loop from (today - h) to today
tday <- Sys.Date()
fday <- tday - h + 1
count <- 1

for (day in as.list(seq(fday, tday, by="day")))
{
  # Take the subset of the data set only necessary for calculating lagged and rolling mean features for the day
  X.subset <- df
  if (count != h){
    X.subset <- head(X.subset, -(h-count))
  }else{}
  X.subset$delta_cases[is.na(X.subset$delta_cases)] <- X.subset$predicted[is.na(X.subset$delta_cases)]
  X.subset <- X.subset %>% select(time, delta_cases, mean_R)
  X.subset <- X.subset  %>% rename(cases=delta_cases, R=mean_R)
  insert_row <- length(X.subset$cases)
  
  # Create features
  X.subset <- create_features(X.subset)
  
  # Construct a matrix only with the 'day'
  X.subset <- tail(X.subset, n=1)
  X.subset[, c("time", "R") := NULL]
  X.subset <- data.matrix(X.subset)
  
  # Update mean_R column of df
  R_prediction <- predict(model.lgb, X.subset)
  
  df$mean_R[insert_row] <- R_prediction
  
  cat(as.character(day),'\t', 'Predicted R ', R_prediction,'\n')
  count <- count + 1
}
```

```{r, fig.width=15, fig.height=3.25, warning=FALSE}
ggplot(df, aes(x=time)) + 
  geom_line(data = head(df, n=-h), aes(x=time, y=mean_R, color='historical'), color='black') + 
  geom_line(data = tail(df, n=h), aes(x=time, y=mean_R, color='predicted'), color='darkred') +
  geom_hline(yintercept=1.0, linetype = 'dotted', color='black') +
  ggtitle("R Forecast") +
  annotate("text", x=2021.305, y=1.38, size = 2.5, label=label, angle=90) +
  annotate("text", x=2020.00, y=2.6, size=3, label='2020-01-01') + 
  annotate("text", x=2020.25, y=2.6, size=3, label='2020-04-01') + 
  annotate("text", x=2020.50, y=2.6, size=3, label='2020-07-01') + 
  annotate("text", x=2020.75, y=2.6, size=3, label='2020-10-01') + 
  annotate("text", x=2021.00, y=2.6, size=3, label='2021-01-01') + 
  annotate("text", x=2021.25, y=2.6, size=3, label='2021-04-01') +
  coord_cartesian(xlim = c(2020, 2021.25), clip = "off") +
  xlab("Time") +
  ylab("R") +
  theme_minimal()
```


```{r, warning=FALSE, message=FALSE}
est_R <- read.csv(url("https://raw.githubusercontent.com/covid-19-Re/dailyRe-Data/master/CHE-estimates.csv"))
est_R <- est_R %>% group_by(date) %>% summarise(mean=mean(median_R_mean))
est_R$date <- as.Date(est_R$date, "%Y-%m-%d")
est_R <- est_R[which(est_R$date >= as.Date('2020-02-25', "%Y-%m-%d")), ]
mean <- est_R$mean
est_R
```

```{r, warning=FALSE, message=FALSE}
ggplot(est_R, aes(x=date)) + 
  geom_line(data = head(est_R, n=-5), aes(x=date, y=mean, color='historical'), color='black') + 
  geom_line(data = tail(est_R, n=5), aes(x=date, y=mean, color='predicted'), color='darkred')
```
